---
title: "DS1_Homework1"
author: "Charlotte Abrams"
output: html_document
date: "2023-03-31"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

Problem 1

```{r}
data("penguins", package = "palmerpenguins")
penguins
unique(penguins$species)
unique(penguins$island)
```
### Data description

The penguins dataset contains records for `r nrow(penguins)` individual penguins from three species (`r unique(penguins$species)`) living on three islands (`r unique(penguins$island)`). The data include `r ncol(penguins)` variables, and are primarily physiological -- measurements of bill, flippers, body mass, and sex -- in addition to species, island, and year.

The mean flipper length (rounded to the nearest mm) of penguins with nonmissing values in this dataset is `r round(mean(pull(penguins, flipper_length_mm), na.rm = TRUE))` mm.
 
Below is a scatterplot comparing flipper length to bill length, color coded by species. 

```{r scatterplot}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + 
  geom_point()

ggsave("DS1_Homework1_scatterplot.pdf")
```


## Problem 2

### Create data frame

Below we create a tibble consisting of:

* a random sample of size 10 from a standard Normal distribution
* a logical vector indicating whether elements of the sample are greater than 0
* a character vector of length 10
* a factor vector of length 10, with 3 different factor "levels"

```{r create_df}
prob2_df = 
  tibble(
    samp = rnorm(10),
    samp_gt_0 = samp > 0,
    char_vec = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"),
    factor_vec = factor(c("low", "low", "low", "medium", "medium", "medium", "medium", "high", "high", "high"))
  )
```

### Take the mean of each variable

When we try to take the mean of each variable:

* We can successfully take the mean of `samp` because they are all numeric values. 
* We can successfully take the mean of `samp_gt_0` because R interprets the logical vector as a vector of 0's (FALSE's) and 1's (TRUE's). The resulting mean is the proportion of TRUE values.
* We get warnings for the variables `char_vec` and `factor_vec` because neither are numeric nor logical, and can't be interpreted as numeric. The resulting output is `NA`.

```{r mean_pull_var}
mean(pull(prob2_df, samp))
mean(pull(prob2_df, samp_gt_0))
mean(pull(prob2_df, char_vec))
mean(pull(prob2_df, factor_vec))
```

### Convert variables to numeric

Next we apply the `as.numeric` function to the logical, character, and factor variables.

* We can successfully convert `samp` because they are all numeric values already. 
* We can successfully convert `samp_gt_0` to numeric, taking values of 0 (FALSE) and 1 (TRUE).
* We cannot convert `char_vec` to numeric. The result is a warning and a vector of NA's.
* We can successfully convert `factor_vec` to numeric, taking values of 2 for `"low"`, 3 for `"medium"`, and 3 for `"high"`. Numeric values are assigned to the factor levels, which are alphabetical unless otherwise specified.

```{r try_numeric}
as.numeric(pull(prob2_df, samp))
as.numeric(pull(prob2_df, samp_gt_0))
as.numeric(pull(prob2_df, char_vec))
as.numeric(pull(prob2_df, factor_vec))
```